How to select predictive models for decision-making or causal inference?

### 2023-12-15 how_to_select_causal_models.tex
iter

\onecolumn
\maketitle
\begin{abstract}
    \textbf{Objective:}
    As predictive models --\eg from machine learning-- give likely
    outcomes, they may be used to reason on the effect of an intervention, a
    causal-inference task. The increasing complexity of health data has opened
    the door to a plethora of models, but also the Pandora box of model
    selection: which of these models yield the most valid causal estimates? 
    We study such model selection procedures in practical settings:
    finite samples settings and without well-specification assumption.
    \textbf{Methods:}
    Elaborate risks build proxies of the causal error using ``nuisance''
    re-weighting to compute it on the observed data. We evaluate whether
    empirically estimated nuisances, which are necessarily noisy,
    add noise to model selection. Drawing from an extensive empirical
    study based on a simulation and three healthcare datasets
    based on real covariates, we compare different metrics for causal model
    selection.\\
    \textbf{Results:} Among all metrics the classical mean squared error (ie.
    factual $\mu\mathrm{-risk}$) is worse and re-weighting it with propensity score
    ($\mu\mathrm{-risk}_{IPW}$) does not bring much improvements. The
    $R\text{-risk}$, which includes a model of mean outcome and propensity
    scores, leads to the best performances.\\
    \textbf{Conclusions:} We outline that selecting model for causal
    reasoning requires different procedures than standard predictive
    settings: using the so-called $R\text{-risk}$ and computing nuisance
    models with flexible estimators on the train set.
\end{abstract}
\keywords{Model Selection, Treatment Effect, G-formula, Observational Study, Machine Learning, }
\clearpage

>>>>

\onecolumn
\maketitle
\begin{abstract}
    \textbf{Objective:}
    As predictive models --\eg from machine learning-- give likely
    outcomes, they may be used to reason on the effect of an intervention, a
    causal-inference task. The increasing complexity of health data has opened
    the door to a plethora of models, but also the Pandora box of model
    selection: which of these models yield the most valid causal estimates? 
    We study such model selection procedures in practical settings:
    finite samples settings and without well-specification assumption.
    \textbf{Methods:}
    Elaborate risks build proxies of the causal error using ``nuisance''
    re-weighting to compute it on the observed data. We evaluate whether
    empirically estimated nuisances, which are necessarily noisy,
    add noise to model selection. Drawing from an extensive empirical
    study based on a simulation and three healthcare datasets
    based on real covariates, we compare different metrics for causal model
    selection.\\
    \textbf{Results:} Among all metrics the classical mean squared error (ie.
    factual $\mu\mathrm{-risk}$) is worse and re-weighting it with propensity score
    ($\mu\mathrm{-risk}_{IPW}$) does not bring much improvements. The
    $R\text{-risk}$, which includes a model of mean outcome and propensity
    scores, leads to the best performances.\\
    \textbf{Conclusions:} We outline that evaluating models for causal
    reasoning requires different procedures than standard predictive
    settings: using the so-called $R\text{-risk}$ and computing nuisance
    models with flexible estimators on the train set.
\end{abstract}
\keywords{Model Selection, Treatment Effect, G-formula, Observational Study, Machine Learning, }
\clearpage
